---
title: "Computing Confidence Intervals in R for UX Researchers"
author: Darrell J. Penta, PhD
output:
  html_document:
    keep_md: true
    number_sections: yes
editor_options:
  chunk_output_type: inline
---
# Introduction
The functions in this package compute success rates and confidence intervals following [Jeff Sauro's excellent write up on MeasuringU](https://measuringu.com/wald/), [Lewis and Sauro (2006)](http://uxpajournal.org/wp-content/uploads/sites/8/pdf/JUS_Lewis_May2006.pdf), and [this page](http://www.measuringux.com/adjustedwald.htm) by Tom Tullis, which includes a downloadable [Excel spreadsheet](http://www.measuringux.com/AdjustedWald.xls) for computing the Adjusted Walk Confidence Interval.

Lewis and Sauro (2006, p. 144) offer several recommendations:

1. Always compute a confidence interval, as it is more informative than a point estimate. For most usability work, we recommend a 95% adjusted-Wald interval
(Sauro & Lewis, 2005).

2. If you conduct usability tests in which your task completion rates typically take a wide range of values, uniformly distributed between 0 and 1, then you should use the LaPlace method. The smaller your sample size and the farther your initial estimate of p is from .5, the more you will improve your estimate of p.

3. If you conduct usability tests in which your task completion rates are roughly restricted to the range of .5 to 1.0, then the best estimation method depends on the value of x/n.  
  * If x/n ≤ .5, use the Wilson method (which you get as part of the process of computing an adjusted-Wald binomial confidence interval). 
  * If x/n is between .5 and .9, use the
MLE. Any attempt to improve on it is as likely to
decrease as to increase the estimate’s accuracy. 
  * If x/n ≥ .9, but less than 1.0, apply either the LaPlace or Jeffreys method. DO NOT use Wilson in this range to estimate p, even if you have computed a 95% adjusted-Wald confidence interval! (3d) If x/n = 1.0, use the Laplace method.

4. Always use an adjustment when sample sizes are small (n<20). (It does no harm to use an adjustment when sample sizes are larger.)


## Adjusted-Wald binomal confidence interval
The equation for the Adjusted-Wald confidence interval is given in (1)

(1) 

$$\hat{p}_{adj} \pm z_{\alpha} \times \sqrt{\frac{\hat{p}_{adj}(1-\hat{p}_{adj})}{n_{adj}}}$$

$n$ = total trials  
$\hat{p}$ = proportion of success trials  
$z_{\alpha}$ = the desired critical z-value (defaulting to 1.96 in the `success_rate()` function of this package)  
$\hat{p}_{adj} = \left(\frac{(n \times \hat{p} + z_{\alpha}^2/2)}{n + z_{\alpha}^2} \right)$  
$n_{adj} = n + z_{\alpha}^2$  

## Point estimators
In addition to the Wilson method, which is used in computing the Adjusted-Wald binomial confidence interval, Lewis and Sauro (2006) also describe the use of the Laplace and Maximum Likelihood Estimate (MLE) for point estimation.

### Laplace Method
The equation for the Laplace method is given in (2)

(2)
$$\frac{(x + 1)}{(n + 2)}$$

$x$ = the observed number of success trials  
$n$ = the total number of trials   

### Maximum Likelihood Estimate (MLE)
The equation for the Maximum Likelihood Estimate method is given in (3)

(3)
$$\frac{x}{n}$$

$x$ = the observed number of success trials  
$n$ = the total number of trials  


# Working with the UserTests functions

## Install packages
Install and load the ``UserTests`` package and some other packages.

```{r echo=TRUE}

#install.packages("tidyverse", repos = "http://cran.us.r-project.org" ) 
#install.packages("devtools", repos = "http://cran.us.r-project.org")
#devtools::install_github(repo = "darrellpenta/UserTests")

library(tidyverse) #For importing and exporting data, wrangling data, making figures
library(devtools) #For installing the UserTests package from GitHub
library(UserTests)
```

## Quick Analyses

If you just want to run analyses on a single task for which you know the success rate and the number of trials, provide that information to the ``UserTests::success_rate`` function, as below. 


```{r, echo=TRUE}
mydata <-
  success_rate(.success=7, .trials=17)
mydata
```

The results returned indicate:

1. **successes**  
The total number of success
1. **trials**  
The total number of trials
1. **orig.succ.pct**  
The raw success rate as a percentage
1. **estimator**  
The name of the method used to adjust the success rate (see [Lewis & Sauro, 2006](http://uxpajournal.org/wp-content/uploads/pdf/JUS_Lewis_May2006.pdf])).
1. **success.pct**  
The adjusted success rate as a percentage
1. **low.ci.pct**  
The lower confidence limit as a percentage
1. **hi.ci.pct**  
The upper confidence limit as a percentage

## Analyzing larger data sets

### Get your data in order
 
Start by properly preparing your data file, which should be saved in``.csv`` format. The data file should have either 3 or 4 columns, depending upon whether one or two test groups are being analyzed. 

In both cases, three of the columns should be: **Participant** (numeric value), **Task** (numeric value), and **Success** (numeric value, coded as 1=success, 0=failure). The fourth (optional) column should be **Group** (character/text value). See table 1:

**Table 1. Example task completion data set**

Participant | Task | Success | Group   
----------- | ---- | ------- | ----- 
 1          | 1    |  1      | US    
 1          | 1    |  0      | THEM      
 1          | 2    |  1      | US    
 1          | 2    |  1      | THEM     
 2          | 1    |  1      | US    
 2          | 1    |  0      | THEM      
 2          | 2    |  0      | US    
 2          | 2    |  1      | THEM
 ...        | ...  | ...     | ...


### Import the data
Next, import the ``.csv`` file into R using the [readr](https://cran.r-project.org/web/packages/readr/README.html) package.

```{r}
mydata<-
  readr::read_csv("sample data/sample_data_2.csv")
head(mydata)
```


### Create a table of the adjusted completion rate means and confidence intervals
Run the ``success_rate`` function on your data set to view a table of summarized data. The ``head`` function in the code below is a convenient way to view the first few rows of data.
```{r echo=TRUE}
mytable<-
  success_rate(mydata)
head(mytable)

```

### Export the table

You can export the table as a ``.csv`` file. Just include the path to the location where you want to save the file to the ``path`` argument in the ``write_csv`` function, as in the example below. 
```{r eval=FALSE}
readr::write_csv(mytable, path = "..\MyDesktop\MyUsabilityStudy\completion-rates.csv")
```

## Create a figure
If you need a figure, use the ``comp_figure`` function. You can overwrite the default labels by providing your own to the appropriate arguments in the function.

```{r echo=TRUE}
myfigure <-
  success_rate_fig(mytable, xlabel="Test Task", ylabel = "Success (%)", legend_lab="Groups")
myfigure
```

## Saving the figure

To save the figure, provide the output format and path. The easiest way is to combine both of these in one string.  Acceptable file formats include:

* eps
* ps
* tex
* pdf
* jpeg
* tiff
* png
* bmp
* svg
* wmf

```{r eval=FALSE}
ggsave("../MyDesktop/usability-test-figure.png")
```

Visit my website, [darrelljpenta.com](http://www.darrelljpenta.com) or <a href="mailto:darrellpenta@gmail.com">email me</a>
